Detecting Facial Expressions in Professional Tennis Matches

### 2017-01-09 Report06-12.Rmd
Updates
Additional reference

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
require(pander)
require(ggplot2)
require(imager)
require(devtools)
require(knitcitations)
require(RefManageR)
require(readr)
require(knitr)
require(grid)
require(tidyverse)
require(magick)
require(kfigr)

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE)

@@

    }
    if(matchBox){
      p <- p + geom_path(aes(x=x, y=450-y, group = type, colour=boxID), faceData) + guides(colour="none")
    }
    else{
      p <- p + geom_path(aes(x=x, y=450-y, group = type, colour=type), faceData)
    }
  return(p)
  }
}

@@

#change this to facesVenn with binary information
ALLmetaIMGFaces<-ALLmetaIMG%>%
  filter(!is.na(boxID)) %>% 
  select(-ID, -facecounter)%>%
  mutate(fileID = as.numeric(factor(file))) %>%
  mutate(FaceKey=paste(fileID, boxID, sep="-")) %>%
  mutate(FaceID=paste(fileID, boxID, substring(type, 1,2), sep="-"))

@@

Many Tennis professionals believe that tennis is a game heavily affected by the mental states of the players. The opportunity for researching this concept presents itself with the hope of improving the playing and coaching of Tennis players. 
By statistically analysing the faces and expressions of players during a match there is a hope that insight may be gained into the effects of the mental state on the outcome of a match.
Facial expressions during competition provide the most direct insight into a player's thoughts. The aim of this project is to begin to develop methods to collect accurate information about the facial expressions of elite tennis athletes during matchplay.

@@

Making use of the recognition software's currently available presents a challenge as high performance sports are not the intended uses of such software's.
Their capabilities are often limited to their intended security and surveillance uses. This report aims to analyse the application of these software's to a broadcast to find a suitable software and API to use to analyse a pre-recorded tennis broadcast file.

####Project Aim
- Find the best facial recognition software for tennis matches
- Build manually classified dataset of faces
- Compare software performance under varied conditions

@@

The sample consisted of a set of 6404 still images. To produce these images, a still shot of the frame was taken at every three seconds, for the length of each 5 minute segment. The stills were provided by Tennis Australia for use in this research, these segments were taken from 105 video files, which were the broadcast of the Tennis Matches shown on the Seven Network during the Australian Open 2016. The sample included an equal amount of singles tennis matches played between females and males. The rounds of the competition vary as to not limit the pool of players to only those who progressed, though there was a higher chance of advancing players reappearing.

@@

The final software we chose to analyse was Google Vision API. Due to Google's expansion in many web based solutions we researched for a facial recognition software. While it was somewhat difficult to access

@@

It was decided that it would be reasonable for a software to detect was decided to be a face size larger than 20 by 20 pixels, as it was specified that the software had minimum distances between the eyes on a face for recognition. (reference table with values)
If it was the face of a player it was always recorded if it captured their face. The back of the head was not able to be picked up by any software, so after a demo trial these faces were not classified manually.
Crowd shots provided difficultly in determining which faces were reasonable to classify. As these faces were not the intended targets of the recognition these faces were contributing to our understanding of the softwares. 
The same face size standard applied to crowd memebers, but focus was placed on the most prominent faces. For each of these faces, we collected information on the following attributes:

@@

####Manual Annotation
To record these questions a Shiny App was created. We called this Application our ManualClassificationProgram. This helped to provide information for all attributes readily and consistently. This worked by presenting an image, using the imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attibutes in `r figr("Scene-Attributes-Table", TRUE, type="Table")` as shown above. If there was a face in the image we were able to highlight the square 'Face Box'. This recorded the x and y coordinates of the box drawn, and when the save button was hit it saved all the radio button selections and the 'Face Box' coordinates to a CSV file.

Following the saving of each face, and the Scene radio button selections, the save button would then save the Scene selections to a separate CSV file. This data can be seen in the ManualClassifiedFaces.csv file located in the appendix.

If there were issues, the CSV files were able to be edited, this was reserved for extreme circumstances. And a lot of care was taken to ensure the first selections were correctly submitted and applied to the correct Faces and Scenes.

The results in the CSV files showed the file name, and the attributes relating to the Scene being considered. 
This can be seen in the ManualClassifiedScenes.csv file located in the appendix.

####Software Recognition
The software choices allowed for POST requests to be sent via the internet. To access the APIs through R we enlisted the httr package, using functions from this package a script was written for each software API, it would loop through the images, individually posting a request for each image to be analysed. These scripts included retrieving the information provided and converting it into a usable format.
One interesting anomoly was found when using the Skybiometry software as it limited the amount of requests per minute. We accounted for this by stalling the posts for the amount of waiting time the software notified, and checking until the time lapsed and the script could continue looping.

####Data Processing
The data was spread across 6 files. For each software we had the classified faces csv file, these are located in the appendix.
The collation of the results from the Manual Recognition Program created the files ManualClassifiedFaces and ManualClassifiedScenes.

A single data set was created to combine all necessary information in the previously mentioned files for our analysis. The information in the dataset,(cite ALLmetaIMG.csv,) was carefully considered. It considers the identify of each face, and all relative face attributes, as well as the image file the face was found in, from this information each face was able to be uniquely identified. Also included was information on the software that found it, and the time it took the software to identify the face. It also has a record of how many faces had been indentified in the image by counting each additional recognised face.

To find whether the softwares were recongising the same faces a function was created. As the location and size of the boxes around the faces were recorded, these values were used to see if a particular identified face box matched a manually identified face.
 This function uses the information of each face and compares the intersecting regions of the polygons created by the x,y coordinates of Manual Faces and other software's faces, to determine if the same face was recognised. We determined the ratio of intersecting area to total area must be greater than 0.1 to be considered the same face. This allowed us to compare the identification areas, as well as contrast the identified faces of each software.

@@

## Results
Using the database of the combined results, we were able to compare the performance of the softwares.
Firstly, we considered how many individual faces the softwares were able to detect.

```{r Face-Bounding-Box-Data, results='hide', type="figure", fig.cap='..', anchor="figure"}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge") +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE)
```

The bar chart above shows the number of Bounding Boxes produced by each software, comparing the height of the bars indicates that Google's Facial Recognition software recognised almost 1000 more faces than the next best software, Microsoft.

#####**Graph 4.** Face Bounding Box Data: Box Plot 
```{r G4 Face Bounding Box Data, anchor="figure"}
ggplot(ALLmetaIMG, aes(x = factor(x = factor(type)), fill=type, y = size)) + geom_boxplot() + scale_y_continuous(trans='log') +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Google","Animetrics","Manual","Microsoft","Skybiometry"))+ylab("Size of Face Bounding Boxes")+guides(fill=FALSE)

@@

This box and whisker plot shows the differences between the sizes of the 'Face Bounding Boxes' produced by each software. Where a 'Face Bounding Box' refers to the size of the square the software provided as a location in the image of a face that was detected. On average, Google maps the largest boxes around faces, however Animetrics has the largest box recognised in the set. On average, the smallest faces are recognised by Skybiometry.

@@

Note: Skybiometry boxes are pink, Microsoft are blue, Google are Green and Animetrics are Red.

@@

#### Challenges

 it is understandable that there would be many more faces to recognise in these shots than in shots where there is only a player, and therefore many more faces recognised. However in this research

- Access to usable images and videos
- Availability of software
- Using the software
- Time constraints

Long Term Goals
- Better understand emotion's effect on player performance
- Automatic collection of player emotion data from video

@@

(5) Conclusions 
Employ the Google Vision API, which would allow the use of still images, or video (TEST VIDEO) files, reducing the need for stills.
This product - cost in relation
Ease of access - API calling

(6) Future Work
Considering the images used during our study were stills derived from Broadcast video files,
it would be useful to extend further research to deal with the video files directly. 
These companies have the potential to allow for video recognition......

It should also be considered that there are softwares focussed on providing recognition in certain controlled scenarios.
If the study was controlled to focus on certain camera angles that align with the facial angles these security minded programs
are intended for it may be helpful to extend the research to softwares currently available but were deemed impractical for this study.

To check for manual errors, create another app that shows the 'faces' identified by each program, confirm manually whether or not these are faces.
We feel this would be necessary when expanding on the use

We also feel that incorporating audio information from the microphones worn by players may assist in sentiment analysis.

Getting specific camera angle access from the seven network to reduce the work of computers

To undertake sentiment analysis, we would take the boxes of faces found in this set of images. Allowing each face a border we would crop the images and produce an individual face image that would form the data set for emotion recognition. igh enough quality?

Write a program to create images of only that faces

(7)
Appendix
```{r}

@@

citation @Automated

```{r}
#Survey of commercial technologies for face recognition in video

#Report number: DRDC-RDDC-2014-C245

#Affiliation: Department of National Defence of Canada.

>>>>

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(pander)
library(ggplot2)
library(imager)
library(devtools)
library(knitcitations)
library(RefManageR)
library(readr)
library(knitr)
library(grid)
library(tidyverse)
library(magick)
library(kfigr)
library(UpSetR)

knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE, kfigr.prefix=TRUE)

@@

    }
    if(matchBox){
      p <- p + geom_path(aes(x=x, y=450-y, group = c(type, boxID), colour=boxID), faceData) + guides(colour="none")
    }
    else{
      p <- p + geom_path(aes(x=x, y=450-y, group = type, colour=type), faceData)
    }
  return(p)
  }
}

@@

# change this to facesVenn with binary information
ALLmetaIMGFaces<-ALLmetaIMG%>%
  filter(!is.na(boxID)) %>% 
  dplyr::select(-ID, -facecounter)%>%
  mutate(fileID = as.numeric(factor(file))) %>%
  mutate(FaceKey=paste(fileID, boxID, sep="-")) %>%
  mutate(FaceID=paste(fileID, boxID, substring(type, 1,2), sep="-"))

# Create Venn Diagrams
VennLists<-ALLmetaIMGFaces %>%
  split(.$type) %>% 
  map(~ .$FaceKey)%>% 
  fromList() %>%
  upset(order.by="freq", nsets=100)

@@

Many tennis professionals believe that tennis is a game heavily affected by the mental states of the players. The opportunity for researching this "inner game" presents itself with the hope of improving the playing and coaching of tennis players by improving their "mental game". 
By statistically analysing the faces and expressions of players during a match there is a hope that insight may be gained into the effects of the mental state on the outcome of a match.
Facial expressions during competition provide the most direct insight into a player's thoughts. The aim of this project is to begin to develop methods to collect accurate information about the facial expressions of elite tennis athletes during matchplay.

@@

Making use of the recognition software's currently available presents a challenge as high performance sports are not the intended uses of such software's.
Their capabilities are often limited to their intended security and surveillance uses. @Boston addresses the 'lack of robustness of current tools in unstructured environments' that this paper faces and applies to a sports environment.
This report aims to analyse the application of these software's to a broadcast to find a suitable software and API to use to analyse a pre-recorded tennis broadcast file.

####Project Aim
The major aim of the present study were to determine the feasibility of using currently available facial recognition algorithms for extracting facial information from players during broadcasts of professional matches by comparing the performance of several popular facial recognition APIs. The performance of the evaluated software was compared against manual classification obtained notational tool developed by the authors. In addition to looking at the overall performance, we also evaluated image factors that influences the performance of each service.

@@

The sample consisted of a set of 6404 still images. To produce these images, a still shot of the frame was taken at every three seconds, for the length of each 5 minute segment. The stills were provided by Tennis Australia for use in this research, these segments were taken from 105 video files, which were the broadcast of the tennis Matches shown on the Seven Network during the Australian Open 2016. The sample included an equal amount of singles tennis matches played between females and males. The rounds of the competition vary as to not limit the pool of players to only those who progressed, though there was a higher chance of advancing players reappearing.

The sample included images that contained the faces of many people, this included players, staff on the court and fans in the crowd. These faces were included in the manual annontations as they were likely to be found by the softwares selected. To be able to contrast the abilities of the softwares and provide information on how to differentiate between players and other people for further research the sample was not filtered at this stage.

@@

The final software we chose to analyse was Google Vision API. Due to Google's expansion in many web based solutions we searched for a facial recognition software.

@@

It was decided that it would be reasonable for a software to detect was decided to be a face size larger than 20 by 20 pixels, as it was specified that the software had minimum distances between the eyes on a face for recognition. (reference table with values)
If it was the face of a player it was recorded if it obviously showed their face. The back of the head was not able to be picked up by any software, so after a demo trial these faces were classified manually  but reclassified as other.
Crowd shots provided difficultly in determining which faces were reasonable to classify. As these faces were not the intended targets of the recognition these faces were contributing to our understanding of the softwares. 
The same face size standard applied to crowd members, but focus was placed on the most prominent faces. For each of these faces, we collected information on the following attributes:

@@

####Manual Annotation
To record the selections of attribute values for each face and scene a Shiny @shiny App was created. We called this Application our ManualClassificationProgram. This helped to provide information for all attributes quickly and consistently.
The program worked by presenting an image, using the imager @imager package, from the sample of images that had not yet been considered, appearing underneath were a set of radio buttons corresponding to the Scene Attibutes in `r figr("Scene-Attributes-Table", TRUE, type="Table")` as shown above.

If there was a face in the image the annotator was able to highlight the square 'Face Box'. This changed the display and presented a set of Attributes with radio buttons, this allowed information to be recorded for the specific face. This recorded the x and y coordinates of the corner points of a box drawn by the mouse, and when the save button was hit it saved all the radio button selections and the 'Face Box' coordinates to a CSV file^[https://github.com/mvparrot/face-recognition/blob/master/ManualClassifiedFaces.csv].

When a face was not selected, the radio buttons showed the Scene attributes and the radio buttons with the possible selections the annotated was able to choose from. When in this display, selecting the save button would then save the Scene selections to a specific CSV file^[https://github.com/mvparrot/face-recognition/blob/master/ManualClassifiedScenes.csv].

If there were issues, the CSV files were able to be edited, this was reserved for extreme circumstances. As a lot of care was taken to ensure the first selections were correctly submitted and applied to the correct Faces and Scenes.

####Software Recognition
The software choices allowed for POST requests to be sent via the internet. To access the APIs through R we enlisted the httr package, using functions from this package a script was written for Google, Animetrics^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/animetrics.R], Microsoft^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/microsoftAPI.R] and Skybiometry^[https://github.com/mvparrot/face-recognition/blob/master/SoftwareRequestScripts/autoSkybiometry.R]. These scripts contained loops that would move through the images, individually posting a request for each image to be analysed. These scripts included retrieving the information provided and converting it into a usable format for our analysis.
One interesting anomoly was found when using the Skybiometry software as it limited the amount of requests per minute. We accounted for this by stalling the posts for the amount of waiting time the software notified, and checking until the time lapsed and the script could continue looping.

####Data Processing
The data needed for our analysis was spread across six files. For each software we had the information on the location of the Facial Bounding Boxes, as well as the time taken for the software to find the information. Some of the softwares also provided a more detailed level of information.

The collation of the results from the Manual Recognition Program created
two CSVs, ManualClassifiedFaces^[https://github.com/mvparrot/face-recognition/blob/master/ManualClassifiedFaces.csv] and ManualClassifiedScenes^[https://github.com/mvparrot/face-recognition/blob/master/ManualClassifiedScenes.csv].

A single data set was created to combine all necessary information in the previously mentioned files for our analysis. The information in the data set^[https://github.com/mvparrot/face-recognition/blob/master/ALLmetIMG.csv], was carefully considered. It considers the identify of each face, and all relative face attributes, as well as the image file the face was found in, from this information each face was able to be uniquely identified. Also included was information on the software that found it, and the time it took the software to identify the face. It also has a record of how many faces had been indentified in the image by counting each additional recognised face.
To do so, we gathered the name of the file the face was found in and the software Type the Face Bounding Box was determined by. The automatically determined time values were also included. The minimum and maximum x and y values were drawn from different values in each software's CSV files. This required some processing to align the differing values to be comparable.

To find whether the softwares were recognising the same faces a function was created. As the location and size of the boxes around the faces were recorded, these values were used to see if a particular identified face box matched a manually identified face, or a region found by another software.
This function uses the information of each face and compares the intersecting regions of the polygons created by the x,y coordinates of Manual Faces and other software's faces, to determine if the same face was recognised. We determined the ratio of intersecting area to total area must be greater than 0.1 to be considered the same face. This allowed us to compare the identification areas, as well as contrast the identified faces of each software. This contributed another variable, boxID, to the data set^[https://github.com/mvparrot/face-recognition/blob/master/ALLmetIMG.csv].

@@

Using the data set^[ALLmetaIMG] of the combined results, we were able to compare the performance of the softwares.
Firstly, we considered how many individual faces the softwares were able to detect in 
`r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` .

## Results

```{r Face-Bounding-Box-Data, results='hide', type="figure", fig.cap="Face Bounding Box Data:", anchor="figure"}
ggplot(ALLmetaIMG, aes(x = type, fill=type)) + geom_bar(position="dodge") +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Manual","Google","Microsoft","Skybiometry","Animetrics")) +ylab("Number of Potential Faces") + guides(fill=FALSE)
```

The bar chart in 
`r figr("Face-Bounding-Box-Data", TRUE, type="Figure")` above shows the number of Bounding Boxes produced by each software, comparing the height of the bars indicates that Google's Facial Recognition software recognised almost 1000 more faces than the next best software, Microsoft.

```{r Face-Bounding-Box-Data-Box, anchor="figure"}
ggplot(ALLmetaIMG, aes(x = factor(x = factor(type)), fill=type, y = size)) + geom_boxplot() + scale_y_continuous(trans='log') +xlab("Facial Recognition Softwares")+scale_x_discrete(limits=c("Google","Animetrics","Manual","Microsoft","Skybiometry"))+ylab("Size of Face Bounding Boxes")+guides(fill=FALSE)

@@

The box and whisker plot in 
`r figr("Face-Bounding-Box-Data-Box", TRUE, type="Figure")` shows the differences between the sizes of the 'Face Bounding Boxes' produced by each software. Where a 'Face Bounding Box' refers to the size of the square the software provided as a location in the image of a face that was detected. On average, Google maps the largest boxes around faces, however Animetrics has the largest box recognised in the set. On average, the smallest faces are recognised by Skybiometry.

@@

@@

#### Challenges
It is understandable that there would be many more faces to recognise in these shots than in shots where there is only a player, and therefore many more faces recognised. This provides many faces to sort through to find emotions of a player.

We faced the challenge of accessing usable images of players, and specifically their faces.
- Availability of software
- Using the software
- Time constraints

@@

## Conclusions 
Employ the Google Vision API, which would allow the use of still images, or video (TEST VIDEO) files, reducing the need for stills.
This product - cost in relation
Ease of access - API calling

## Future Work
The Long Term goal of this research is to better understand how the emotion's felt by a player during a match affect player performance. Ultimately we would aim to create a program that automated the collection of player emotion data from throughout a match. This information would be presented in a timeline that allowed match performance, in the form of points won, to be aligned with the emotions felt at certain times throughout.

Considering the images used during our study were stills derived from Broadcast video files, it would be useful to extend further research to deal with the video files directly. The Google Vision API used in this research which produced the best recognition in images does not yet have the potential to detect faces and emotions in a video.

It should also be considered that these are softwares focussed on providing recognition in certain controlled scenarios.
If the study was controlled to focus on certain camera angles that align with the facial angles these security programs are intended to recognise faces in.

Given that Google found many faces that did not match manually annotated face, we considered that we should check for manual errors.
There is the possibility that we could create another app that shows the Facical Bounding Boxes identified by each program, this would allow the annotator to confirm manually whether or not these are faces.

Given that certain Scene attribute combinations produced more facial recognitions than other combinations we should consider limiting the sample of images sent to Google Vision API. This would not only reduce cost but also provide a greater level of detail of the emotions felt by a player during a match. To provide a greater level of information at all points in a match it would be beneficial to derive images from a single camera feed. This feed should match the Scene attributes that provided the most Google faces.

To undertake sentiment analysis, we would take the boxes of faces found in this set of images. Allowing each face a border of pixels, we would crop the images and produce an individual face image that would form the data set for emotion recognition.
We also feel that incorporating audio information from the microphones worn by players may assist in sentiment analysis. By including this information we would be able to define differences between certain emotions that may not be able to be found by facial features only.

@@
